---
title: "Problem Set-5"
author: "Aravind Mannarswamy"
date: "9/27/2021"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    number_sections: yes
    toc: yes
    toc_depth: 3
---

\newpage  

```{r header, echo = FALSE, message = FALSE, warning = FALSE}

library(knitr)
knitr::opts_chunk$set(cache = TRUE,
                      echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      cache.lazy = FALSE, 
                      fig.align = "center", 
                      fig.width = 6, 
                      fig.height = 4)
library(tidyverse)
library(magrittr)
library(scales)
library(gridExtra)
```


# Schitt's Creek Recreation   

## Benefits of immigration of high skilled workers   
1. Fills critical labor gaps in advanced technology and STEM jobs    
2. Instrumental in creating jobs and improving the economy   
3. They often go to US Universities thereby funding higher education for US  


## Project's admissibility   

A project is admissible if and only if the net return on the investment is positive.   

$PV = -10000$   
$r = 0.03$  
$R = 12000$    

$PV_R = 12000/1.03$  
$PV_R = 11650.49$   
$PV + PV_R = -10000 + 11650.49 > 0$   

Hence the project is admissible.   


## Project selection   

The present value of a future payment can be calculated as   

$PV = FV/(1+r)^n$   

```{r}
df = data.frame(Year = c(0,1,2), Bike_park = c(-10000, 13000, 18000), Whitewater = c(-12000, 15000, 18000)) %>% 
  mutate(Bike_park_pv = Bike_park/(1.03^Year),
         Whitewater_pv = Whitewater/(1.03^Year))
df
```

Net Return of Bike Park: `r sum(df$Bike_park_pv)`   
Net Return of White water Park: `r sum(df$Whitewater_pv)`    
They are relatively close but the town should choose `Bike Park`.  



## Considerations on cost vs benefit   

1. Market price: The amount that normal consumers pay for goods. This can be used as an anchor for price level.    
2. Shadow price: Adjusting market prices to capture the true price levels     
3. Consumer Surplus: Since the size of the projects are big, there may be other parties benefiting from it      
4. Inferred Costs and benefits


# Two polluters   

## Cost of pollution reduction   

$MC_1 = e_1$   
$MC_2 = 2e_2$   

Where `e1` and `e2` are the amounts of emissions reduced by the 1st and 2nd firms respectively.   

In the absense of intervention,   
Firm-1: 300 units  
Firm-2: 300 units  
 
```{r}
p1 = ggplot()+
  theme_bw()+
  theme(panel.grid = element_blank())+
  scale_x_continuous(limits = c(0,300), breaks = seq(0,600,60))+
  scale_y_continuous(limits = c(0,600), breaks = seq(0,600,60))+
  labs(title = "Cost of reducing pollution",subtitle = "Firm-1",x = "Units of Pollutant", y = "Marginal Cost")+
  annotate(geom = "segment", x = 0, y = 0, xend = 300, yend = 300, color = "red")+
  annotate(geom = "polygon", x = c(0,60,60,0), y = c(0,0,60,0), fill = "red", alpha = 0.5)+
  geom_hline(yintercept = 60, linetype = "dashed", color = "grey")

p2 = ggplot()+
  theme_bw()+
  theme(panel.grid = element_blank())+
  scale_x_continuous(limits = c(0,300), breaks = seq(0,600,60))+
  scale_y_continuous(limits = c(0,600), breaks = seq(0,600,60))+
  labs(title = "Cost of reducing pollution",subtitle = "Firm-2",x = "Units of Pollutant", y = "Marginal Cost")+
  annotate(geom = "segment", x = 0, y = 0, xend = 300, yend = 600, color = "red")+
  annotate(geom = "polygon", x = c(0,60,60,0), y = c(0,0,120,0), fill = "red", alpha = 0.5)+
  geom_hline(yintercept = 120, linetype = "dashed", color = "grey")
grid.arrange(p1, p2, nrow = 1)
```

Cost for Firm-1 = 0.5 x 60 x 60 = 1800   
Cost for Firm-2 = 0.5 x 60 x 120 = 3600   


## How much to cut pollution   

$MC_1 = MC_2$   

$e_1 = 2e_2$   

$e_2/e_1 = 0.5$    

$e_1 + e_2 = 120$   

$2e_2 + e_2 = 120$   

$e_2 = 40$   

$e_1 = 80$   


## Emissions Fee   

$f* = MC_1(e_1^*)$   

$f* = 80$  

Firm-1 pollutes: 300 - 80 = 220 units at 80 per unit   
Firm-2 pollutes: 300 - 40 = 260 units at 80 per unit   

Firm-1 Fee = 220 x 80 = 17600   
Firm-2 Fee = 260 x 80 = 20800   

## Tradeable permits   

Firm-1: Has 180 permits but needs 220 permits. Firm-1 is 40 permits short   
Firm-2: Has 300 permits but needs 260 permits. Firm-2 has 40 extra permits from the optimal solution.  

Firm-1 will buy 40 permits from Firm-2 in this instance.   

Min price = f* = 80   
Min paid = 80 x 40 = 3200   

and the optimal solution point will be maintained   

## Emissions fee or cap-and-trade   

$MC_1 = MC_2$   

$e_1 = 4e_2$   

$e_2/e_1 = 0.25$    

$e_1 + e_2 = 120$   

$4e_2 + e_2 = 120$   

$e_2 = 24$   

$e_1 = 96$   

Emissions Fee   

$f* = MC_1(e_1^*)$   

$f* = 96$   

Firm-1 pollutes: 300 - 24 = 276 units at 96 per unit   
Firm-2 pollutes: 300 - 96 = 204 units at 96 per unit   

Firm-1 Fee = 276 x 96 = 26496   
Firm-2 Fee = 204 x 96 = 19584    


1. Under cap and trade, the pollution reduction for Firm-2 is should be 40 while that under the emissions fee, it will be about 24 units of pollution reduction.  
2. For this reason, Cap-and-trade is a better option for reducing pollution instead of Emissions fee.  





