---
title: "Problem Set-3"
author: "Aravind Mannarswamy"
date: "9/13/2021"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    number_sections: yes
    toc: yes
    toc_depth: 3
---

\newpage  

```{r header, echo = FALSE, message = FALSE, warning = FALSE}

library(knitr)
knitr::opts_chunk$set(cache = TRUE,
                      echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      cache.lazy = FALSE, 
                      fig.align = "center", 
                      fig.width = 6, 
                      fig.height = 4)
library(tidyverse)
library(magrittr)

```

# Statutory Incidence on the producers/suppliers     


$D: P = 95 - 4Q$  
$S: P = 60 + Q$  


## Graph Supply and Demand  
```{r}
ggplot()+
  scale_x_continuous(name = "Quantity(in lbs)", limits = c(0,10), breaks = c(seq(0,10,2),7))+
  scale_y_continuous(name = "Price", limits = c(50,100), breaks = c(seq(0,100,10),67,95))+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 95,xend = 10,yend = 55, color = "red")+
  annotate(geom = "segment", x = 0,y = 60,xend = 10,yend = 70, color = "red")+
  annotate(geom = "segment", x = 7,y = 67,xend = 7,yend = 50, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 67,xend = 7,yend = 67, color = "blue", linetype = "dashed")+
  labs(title = "Artisanal Goat Cheese", subtitle = "Before Taxes")+
  annotate(geom = "polygon", x = c(0,7,0), y = c(95,67,67), fill = "yellow", alpha = 0.4)+
  annotate(geom = "polygon", x = c(0,7,0), y = c(60,67,67), fill = "green", alpha = 0.4)+
  annotate(geom = "text", label = "Demand Curve", x = 2.5, y = 90, color = "red")+
  annotate(geom = "text", label = "Supply Curve", x = 2.5, y = 60, color = "red")+
  annotate(geom = "text", label = "CS = 98", x = 1, y = 80, color = "darkblue") +
  annotate(geom = "text", label = "PS = 24.5", x = 1, y = 65, color = "darkblue")+
  annotate(geom = "text", label = "Optimal(Q* = 7,P* = 67)", x = 7, y = 75, color = "darkblue")
  
  
  
```

## Equilibirum price and quantity  

At equilibrium, the supplier and demand curves should intersect. Solving the above two equations,  

$95-4Q^* = 60 + Q^*$  
$5Q^* = 35$  
$Q^* = 7$  

\newpage  

## Supply and demand with tax  

```{r}
ggplot()+
  scale_x_continuous(name = "Quantity(in lbs)", limits = c(0,10), breaks = c(seq(0,10,2),7))+
  scale_y_continuous(name = "Price", limits = c(50,100), breaks = c(50,60,66,71,95))+
  theme_bw()+
  theme(panel.grid = element_blank())+
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 95,xend = 10,yend = 55, color = "red")+
  annotate(geom = "segment", x = 0,y = 60,xend = 10,yend = 70, color = "red")+
  annotate(geom = "segment", x = 0,y = 65,xend = 10,yend = 75, color = "red", linetype = "dashed")+
  annotate(geom = "segment", x = 6,y = 71,xend = 6,yend = 50, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 66,xend = 6,yend = 66, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 71,xend = 6,yend = 71, color = "blue", linetype = "dashed")+
  labs(title = "Artisanal Goat Cheese", subtitle = "After Taxes: Tax incidence on suppliers")+
  annotate(geom = "polygon", x = c(0,6,0), y = c(95,71,71), fill = "yellow", alpha = 0.4)+
  annotate(geom = "polygon", x = c(0,6,0), y = c(60,66,66), fill = "green", alpha = 0.4)+
  annotate(geom = "polygon", x = c(0,6,6,0), y = c(66,66,71,71), fill = "red", alpha = 0.4)+
  annotate(geom = "polygon", x = c(6,7,6), y = c(66,67,71), fill = "blue", alpha = 0.2)+
  annotate(geom = "text", label = "Demand Curve", x = 2.5, y = 90, color = "red")+
  annotate(geom = "text", label = "Supply Curve(without tax)", x = 2.5, y = 60, color = "red")+
  annotate(geom = "text", label = "Supply Curve(with tax)", x = 8, y = 80, color = "red")+
  annotate(geom = "text", label = "CS = 72", x = 1, y = 80, color = "darkblue") +
  annotate(geom = "text", label = "PS = 18", x = 1, y = 64, color = "darkblue")+
  annotate(geom = "text", label = "Tax = 30", x = 1, y = 69, color = "darkblue")+
  annotate(geom = "text", label = "DWL = 2.5", x = 6.8, y = 58, color = "darkblue")+
  annotate(geom = "text", label = "Optimal(Qt* = 6,Ps* = 66, Pc* = 71)", x = 7, y = 85, color = "darkblue")+
  annotate(geom = "segment", x = 6.8, y = 59, xend = 6.5, yend = 68, arrow = arrow(length = unit(0.25, "cm")), color = "red")+
  annotate(geom = "segment", x = 8, y = 78, xend = 6.5, yend = 72, arrow = arrow(length = unit(0.25, "cm")), color = "red")
  

```

## New equilibrium  

$65+Q = 95-4Q$  
$5Q = 30$  
$Q = 6$  

$P_c = 65+Q$  
$P_c = 71$   

$P_s = 60 + Q$  
$P_s = 66$  

## Consumer Surplus  

Before tax:  $(95 - 67)*7/2 = 98$   
After tax: $(95-71)*6/2 = 72$  

## Producer Surplus  

before tax:  $(67-60)*7/2 = 24.5$  
after tax: $(66-60)*6/2 = 18$   

## tax revenue  

tax = $t * Q_t = 5 * 6 = 30$   

## Dead weight loss   

Method-1:  
$DWL = PS + CS - PS_t - CS_t - tax$   
$DWL = 98 + 24.5 - 72 - 18 - 30 = 2.5$   

Method-2:  
$DWL = (71-66)*1/2 = 2.5$  

## Who pays the tax   

1. The consumer pays the tax regardless of the optics  
2. The only situation where the producer will bear the tax is when they have no control over the quantity produced, in other words when the supply is INELASTIC  

# Story of a monopolist  


```{r}
Qo = seq(0,20,0.1)
Do = 48 - Qo
Co = 28 + 4*Qo + Qo^2
MCo = 4 + 2*Qo
MRo = 48 - 2*Qo

df = data.frame(Qo = Qo, Do = Do, MCo = MCo, MRo = MRo) %>% 
  pivot_longer(!Qo, names_to = 'type', values_to = 'cost')

df %>% 
  ggplot()+
  theme_bw()+
  theme(panel.grid = element_blank(),
        legend.title = element_blank())+
  aes(x = Qo,y = cost, color = type, group = type)+
  geom_line()+
  scale_x_continuous(limits = c(0,22), breaks = c(0,5,10,11,15,20))+
  scale_y_continuous(limits = c(0,50), breaks = c(0,10,20,30,37, 40, 50))+
  annotate(geom = "segment", x = 11,y = 37,xend = 11,yend = 0, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 37,xend = 11,yend = 37, color = "blue", linetype = "dashed")+
  labs(x = "Quantity", y = "Price", title = "Monopolist",subtitle = "Before tax situation")
```

## Price and Quantity   

Price = P  
Quantity = Q  
Demand: $P = 48 - Q$  
Cost: $C = 28 + 4Q + Q^2$  
Marginal Cost: $MC = 4 + 2Q$  
Revenue: $R = PQ$  
$R = 48Q - Q^2$  
Marginal Revenue: $MR = 48 - 2Q$  
  
At equilibrium, MR = MC  

$48 - 2Q = 4 + 2Q$  
$Q = 11$   

## profit without taxes  

Profit = Revenue - Cost  
$\pi = R - C$  
$\pi = 48Q - Q^2 - 28 - 4Q - Q^2$  
$\pi = -2Q^2 + 44Q - 28$  
$\pi = 214$  


```{r}
Qo = seq(0,20,0.1)
Do = 48 - Qo
Co = 28 + 4*Qo + Qo^2
MCo = 4 + 2*Qo
MRo = 48 - 2*Qo
MCt = 8 + 2*Qo

df = data.frame(Qo = Qo, Do = Do, MCo = MCo, MRo = MRo, MCt = MCt) %>% 
  pivot_longer(!Qo, names_to = 'type', values_to = 'cost')

df %>% 
  ggplot()+
  theme_bw()+
  theme(panel.grid = element_blank(),
        legend.title = element_blank())+
  aes(x = Qo,y = cost, color = type, group = type)+
  geom_line()+
  scale_x_continuous(limits = c(0,22), breaks = c(0,5,10,11,15,20))+
  scale_y_continuous(limits = c(0,50), breaks = c(0,10,20,30,38, 40, 50))+
  annotate(geom = "segment", x = 11,y = 37,xend = 11,yend = 0, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 37,xend = 11,yend = 37, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 10,y = 38,xend = 10,yend = 0, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 38,xend = 10,yend = 38, color = "blue", linetype = "dashed")+
  labs(x = "Quantity", y = "Price", title = "Monopolist",subtitle = "With tax situation: On producers")
```

## Taxes - statutory and economic incidence   

1. Statutory incidence in this case is the producer, as they are legally responsible for the tax  
2. Economic incidence however is shared between the producer and the customer(see the final answer). The producer pays 3 while the customer pays 1. In addition to that, the producer also suffers loss of quantity sold, thereby impacting the revenue even further.  

$t = 4$  
$C_t = 28 + 4Q + Q^2 + 4Q$  
$C_t = 28 + 8Q + Q^2$   
$MC_t = 8 + 2Q$  

## Equilibrium with taxes     

$MR = MC_t$  
$48-2Q_t = 8 + 2Q_t$  

$Q_t = 10$  
$P_t = 48 - Q_t$  
$P_t = 38$  

## Profit with taxes   

Profit = Revenue - Cost  
$\pi = R_t - C_t$  
$\pi = 38*10 - (28+8Q_t+Q_t^2)$  
$\pi = 172$    

## Statutory incidence on the customer   

$P = 48 - Q - 4$  
$P_t = 44 - Q$  
Revenue; $P_tQ = 44Q - Q^2$   
Marginal Revenue: $MR_t = 44 - 2Q$   
Cost: $C = 28 + 4Q + Q^2$  
Marginal Cost: $MC = 4 + 2Q$  

At equilibirum;  

$4 + 2Q = 44 - 2Q$  
$Q_t = 10$   
$P_c = 38$   
$P_s = 34$

```{r}
Qo = seq(0,20,0.1)
Do = 48 - Qo
Dt = 44 - Qo
Co = 28 + 4*Qo + Qo^2
MCo = 4 + 2*Qo
MRo = 48 - 2*Qo
#MCt = 8 + 2*Qo
MRt = 44 - 2*Qo 

df = data.frame(Qo = Qo, Do = Do,Dt = Dt, MCo = MCo, MRo = MRo, MRt = MRt) %>% 
  pivot_longer(!Qo, names_to = 'type', values_to = 'cost')

df %>% 
  ggplot()+
  theme_bw()+
  theme(panel.grid = element_blank(),
        legend.title = element_blank())+
  aes(x = Qo,y = cost, color = type, group = type)+
  geom_line()+
  scale_x_continuous(limits = c(0,22), breaks = c(0,5,10,11,15,20))+
  scale_y_continuous(limits = c(0,50), breaks = c(0,10,20,30,38, 40, 50))+
  annotate(geom = "segment", x = 11,y = 37,xend = 11,yend = 0, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 37,xend = 11,yend = 37, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 10,y = 38,xend = 10,yend = 0, color = "blue", linetype = "dashed")+
  annotate(geom = "segment", x = 0,y = 38,xend = 10,yend = 38, color = "blue", linetype = "dashed")+
  labs(x = "Quantity", y = "Price", title = "Monopolist",subtitle = "With tax situation: On Consumers")
```

Regardless of the consumer incidence or the monopolist, the answers are the same.  

## Monopolist profit with tax   

$R = PQ$; $R = 34 * 10$; $R = 340$   
$C = 28 + 4Q + Q^2$; $C = 168$; $\pi_t = 340 - 168$   
$\pi_t = 172$   
So nothing changes with regard to who the statutory incidence falls under,   
