---
title: "Problem Set-2"
author: "Aravind Mannarswamy"
date: "9/06/2021"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    number_sections: yes
    toc: yes
    toc_depth: 3
---

\newpage  

# Play at Schitt's Creek  


```{r header, echo = FALSE, message = FALSE, warning = FALSE}

library(knitr)
knitr::opts_chunk$set(cache = TRUE,
                      echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      cache.lazy = FALSE, 
                      fig.align = "center", 
                      fig.width = 6, 
                      fig.height = 4)
library(tidyverse)
library(magrittr)

```


$U^i = 2k_i^2 + r^2$    

`k` is the private good, candles;   
`r` is the public good, plays;  

Individually, their utilities with the subscript can be given as follows:

For Stevie, $U^S = 2k_S^2 + r^2$    
For Moira, $U^M = 2k_M^2 + r^2$  

and `r` itself can be written as $r = r_S + r_M$   

Their incomes are equal $I_S = I_M = 24$  

Also given is that,  

$P_k = 1$ and   
$P_r = 1$   

Based on what we know so far, here are the budget constraints:  

$k_S + r_S = 24$ and  
$k_M + r_M = 24$    
  
## Criteria for public good   

It meets the following two conditions:  
1. Nonrival: Because adding another person to witness the play is not going to increase the cost.   
2. Non excludable: Because once provided, in an outdoor setting, it is not possible to exclude anyone.    

## How many plays will they provide   

Eliminating the private goods using the budget constraints,   

For Stevie, $U^S = 2(24 - r_S)^2 + (r_S + r_M)^2$    
  
$dU^S/dr_S = 4(24 - r_S)(-1) + 2(r_S + r_M)$      
  
at optimal, this derivative is equal to zero.  

$4(24-r_S) = 2(r_S + r_M)$   

$2(24 - r_S) = (r_S + r_M)$   

$48 - 2r_S = r_S+ r_M$  

$48 - r_M = 3r_S$   

By Symmetry,   

$48 - r_S = 3r_M$  

After some more algebra,  

$r_S = 12$  
$r_M = 12$  
$r = 24$   

Therefore, $K_S = 12$ and $K_M = 12$

Each by themselves, they would have choose to provide 12 plays each and the total to the town would be 24 plays.  

## Shortfall   

If the socially optimal number of plays is 28, then the town will be short by 4 plays as Stevie and Moira together would finance 24 plays.  

## Only if Stewie and Moira had known before   

In this particular setting,   

$r = r_S + r_M + 4$   

For Stevie, 
$U^S = 2(24 - r_S)^2 + (r_S + r_M + 4)^2$    

$d U^S/dr_S = -4(24-r_S) + 2(r_M + r_S + 4)$ should be equal to zero at optimal   

$4(24-r_S) = 2(r_M + r_S + 4)$   

$48 - 2r_S = r_M + r_S + 4$   

$r_S = (44 - r_M)/3$   

By symmetry,   

$r_M = (44-r_S)/3$   

Solving, we get   

$r_S = 11$  
$r_M = 11$   

And in total they will provide 22 plays. So with the town providing the expected shortfall of 4, there will be a total of 26 plays! Well, this is not good. So Steve and Moira having that additional information didnt help the town to achieve the optimal number of plays.   



# Bob has a motorcycle!   

$MPC = 10$  

$MB = 30 - 4Q$   


## Bob's private solution   

Bob will try to maximize his utility by optimizing to a point where the marginal costs and benefits are the same. In other words, he will stop paying for more when it is not worth the cost.  

$MPC = MB$   

$10 = 30 - 4Q$   

$Q = 5$   

Bob would prefer to ride 5 miles!   

## Marginal Social Cost   

Social Cost is the sum of the private cost of the individual(who also reaps benefit) and the rest of the society.  

Social Cost = Marginal Private Cost + Marginal Damage  

$MSC = MPC + MD$  

$MSC = 10 + 5 + Q$   

$MSC = 15 + Q$  

## Socially Optimal Solution   

Taking the social cost into account,   

$MSC = MB$  

$Q + 15 = 30 - 4Q$  

$Q = 3$   

Bob should ride 3 miles for a socially optimal solution.  


## Graph the scenario   

```{r}
library(ggplot2)
# function to invoke a plot object
ggplot()+ 
  # Set x-axis limit
  scale_x_continuous(limits = c(0,10), breaks = seq(0,10))+
  # Set y-axis limit
  scale_y_continuous(limits = c(0,30), breaks = c(seq(0,30,5),18))+
  # I like plain background(instead of default grey)
  theme_bw()+
  # I dont like the gridlines in the background for this chart
  theme(panel.grid = element_blank())+
  # to draw a straight line, we need x1,y1 and x2,y2
  annotate(geom = "segment", x = 0, y = 0, xend = 10, yend = 0, color = "grey")+
  # to draw a straight line, we need x1,y1 and x2,y2; also changing linetype
  annotate(geom = "segment", x = 0, y = 10, xend = 10, yend = 10, color = "darkgreen", linetype = "dashed")+
  # to draw a straight line, we need x1,y1 and x2,y2; the next 5 lines are just straight lines
  annotate(geom = "segment", x = 0, y = 30, xend = 7.5, yend = 0, color = "red", linetype = "dashed")+
  annotate(geom = "segment", x = 0, y = 15, xend = 5, yend = 20, color = "dodgerblue", linetype = "dashed")+
  annotate(geom = "segment", x = 5, y = 20, xend = 5, yend = 0, color = "grey", linetype = "dashed")+
  annotate(geom = "segment", x = 3, y = 18, xend = 3, yend = 0, color = "grey", linetype = "dashed")+
  annotate(geom = "segment", x = 0, y = 18, xend = 3, yend = 18, color = "grey", linetype = "dashed")+
  # How about labeling the chart and axes
  labs(x = "Quantity(Q)", y = "Price/Cost", title = "Bob and his motorcycle")+
  # Shaded region. Specify the closed vertices and the color and the intensity of the color
  annotate(geom = "polygon", x = c(3,3,5), y = c(18,10,10), fill = "yellow", alpha = 0.4)+
  annotate(geom = "polygon", x = c(3,5,5), y = c(18,20,10), fill = "dodgerblue", alpha = 0.4)+
  # Place the text in x,y coordinate and give the text
  annotate(geom = "text", label = "MPC", x = 9, y = 11, color = "darkgreen")+
  annotate(geom = "text", label = "MB", x = 1, y = 28, color = "red")+
  annotate(geom = "text", label = "MSC", x = 1, y = 15, color = "dodgerblue")+
  annotate(geom = "text", label = "8", x = 3.5, y = 12, color = "black")+
  annotate(geom = "text", label = "10", x = 4.5, y = 16, color = "black")

```


## Bob's loss   


The shaded traingle(yellow) represents Bob's loss and that is given as follows:  

Bobs Loss = 0.5 * (5-3) * (18 - 10) = 8 units   

Bob really loses this as a *surplus*  


## Community gain   

The community gains when Bob rides fewer miles, fewer than 5 to be specific in this instance.   

If Bob rides 5 miles, Societal loss(incl Bob) = 50(from Bob) + 37.5(by the rest)  = 87.5  
If Bob rides 3 miles, Societal loss(incl Bob) = 30(from Bob) + 19.5(by the rest) = 49.5   

Hence, the society has avoided a loss of 87.5 - 49.5 = 38

But this 38 also contains Bob's surplus loss of 8 units, and his cost of 20 if he would have driven the additional 2 miles.   

Hence, the net gain for the society is 38 - 20 - 8 = 10 units!

Alternatively, the same solution can be found by finding the area of the triangle shaded in blue(0.5 * (5-3)*(20-10)).   

here is detailed summary:   

Total avoided loss/damage to the town:  18 = 0.5 * (8 + 10) * (5-3);   
Bob's lost surplus in this 18 units = 8   
Net towns gain = 18 - 8 = 10






